Smooth Latent Dirichlet Allocation in Metabolomics

### 2016-12-03 docs/MSci_proposal/mprop.tex
Gibbs sampling review.

>>>>

\subsection{Finding scientific topics}

% Intro:
% - The paper
% - Sampling as a method of inference
% - Experiments
\par The paper on "Finding scientific topics" [CITATION] suggests an alternative approach to the inference in latent Dirichlet allocation. To be more specific, the paper provides a readable introduction on how to apply a Markov chain Monte Carlo (MCMC) algorithm as a sampling-based method for the inference in topic modelling. Further, the authors discuss the method's application settings in vision and conduct an experiment on deriving topics from the abstracts of papers published in "Proceedings of the National Academy of Sciences" (PNAS). Finally, the authors compare the performance of the suggested sampling-based inference method to variational inference methods.

% The method (Gibbs Sampling)
% - Introduce the meaning of phi and theta.
% - Introduce the inference method
\par The inference method using a Markov chain Monte Carlo algorithm can be referred by \textit{Gibbs sampling}. The authors claim that the method provides a first-order approximation in deriving the topics on the given corpus. That is, we can establish quantitative reasoning on the document correlation in terms of content and content's change over time. Recall that by $\theta$ we denote the topic distribution over documents and by $\phi$ we denote the topic distribution over words. In Gibbs sampling, these values can be obtained by examining the posterior distribution. Essentially, the variables are sequentially sampled from their respective distributions. The process draws topic assignment $z_i$ and keeps a track of this assignment with respect to each word and document. Note that $z_i$ is drawn from the following distribution:
\begin{equation}
P(z_i = j | z_{-i}, w) \propto \dfrac{n_{-i, j}^{(w_i)} + \beta}{n_{-i, j}^{(\cdot)} + W\beta}\dfrac{n_{-i, j}^{(d_i)} + \alpha}{n_{-i, \cdot}^{(d_i)} + T\alpha}.
\end{equation}
It follows that this distribution is utilise the previously mentioned topic assignments. For a brief familiarisation with the notation used in the previous expression, note the following: $n_j^{d}$ and $n_j^{w}$ are the counts of topic $j$ assignments in document $d$ and word $w$; $W$ is the number of words; and $T$ is the number of topics. Ultimately, the left side of the expression is $\phi$ and the right side is $\theta$. That is, 
\begin{align*}
\phi_j &= \dfrac{n_{-i, j}^{(w_i)} + \beta}{n_{-i, j}^{(\cdot)}},\\
\theta_j &= \dfrac{n_{-i, j}^{(d_i)} + \alpha}{n_{-i, \cdot}^{(d_i)} + T\alpha}.
\end{align*}
This process of topic assignment would be executed for every word in a document and for all documents in the corpus. The procedure would be repeated until the distributions displayed fluctuations below some set threshold. Notice that the procedure does not require direct assignment of $\theta$ and $\phi$. If needed, the values of these can be sampled upon converge by using the previously provided representations.

% Experiments
% - Time series
% - PNAS
\par The authors have executed Gibbs sampling in the domain of vision. They have represented an image as a document and each pixel as a word. This means that they were able to recover the underlying basis from which the images could be constructed. The performance was compared to the latent Dirichlet allocations models which use variational inference. Note that we define the performance in terms of perplexity, which indicates the uncertainty of the word assignment to a topic. It appears that, with relatively small data sets, Gibbs sampling displays superior performance. Further, the authors have investigated the performance upon varying the number of topics. The results indicate that there exists a trade-off between the model being able to cover the significant underlying topics (low number of topics) and start deriving irrelevant branches (high number of topics).

\par The main experiment was carried in the PNAS data set, note that the processed papers were published from 1991 to 2001. By having a varying notion of time, the authors were able to investigate the prevailing topics. Thus, establish a sense in dynamic topic modelling. Even though the model was not adjusted to smooth the time-series, from the derived results we can deduce the varying relevance of underlying science domains.

% Conclusion
\par We have reviewed an alternative method for the inference in latent Dirichlet allocation. The idea behind the inference process was explained by considering the primary aspects: the notion of counts and the implicit use of hidden variables. Also, we have introduced the experiment settings its results. This allowed to familiarise with an example of topic modelling application and vision. Further, the experiment on PNAS has emphasised the relevance of time series.
